Interpolants and Error Bounds for Modeling and Predicting Variability in Computer Systems

### 2020-07-28 3-Disseration/apps.tex
Updated thesis document, sent draft to Dr. Watson.

Many previous works attempting to model system performance have used simulated environments \cite{grobelny2007fase,wang2009simulation,wang2013towards}. \citet{grobelny2007fase} refer to statistical models as being oversimplified and not capable of capturing the true complexity of the underlying system. Historically, statistical models have been limited to modelling at most one or two system parameters and have therefore not been capable of modeling the complexity of the underlying system. \cite{snavely2002framework,bailey2005performance,barker2009using,ye2010analyzing}. These limited statistical models have provided satisfactory performance in narrow application settings. However these techniques require additional code annotations, hardware abstractions, or additional assumptions about the behavior of the application in order to generate models. In contrast, the approaches that are presented here require no modifications of applications, no architectural abstractions, nor any structural descriptions of the input data being modeled. The techniques used in this work only require performance data as input.

@@

Regression and interpolation have a considerable theoretical base in
one dimension \cite{cheney2009course}. Splines in particular are well
understood as an interpolation technique in one dimension
\cite{de1978practical}, particularly B-splines. Tensor products of
B-splines \cite{unther1996interpolating} or other basis functions have
an unfortunate exponential scaling with increasing
dimension. Exponential scaling prohibits tensor products from being
reasonably applied beyond three-dimensional data. In order to address
this dimensional scaling challenge, C. de Boor and others proposed box
splines \cite{de2013box}, of which one of the approximation techniques
in this work is composed \cite{lux2018novel}.

@@

Predicting the performance of an HPC system is a challenging problem
that is primarily attempted in one of two ways: (1) build a
statistical model of the performance by running experiments on the
system at select settings, or (2) run artificial experiments using a
simplified simulation of the target system to estimate architecture
and application bottlenecks. In this paper the proposed multivariate
modeling techniques rest in the first category, and they represent a
notable increase in the ability to model complex interactions between
system parameters.

Many previous works attempting to model system performance have used
simulated environments to estimate the performance of a system
\cite{grobelny2007fase,wang2009simulation,wang2013towards}. Some
of these works refer to statistical models as being oversimplified and
not capable of capturing the true complexity of the underlying
system. This claim is partially correct, noting that a large portion
of predictive statistical models rely on simplifying the model to one
or two parameters
\cite{snavely2002framework,bailey2005performance,barker2009using,ye2010analyzing}.
These limited statistical models have provided satisfactory
performance in very narrow application settings. Many of the
aforementioned statistical modeling techniques claim to generalize,
while simultaneously requiring additional code annotations, hardware
abstractions, or additional application level understandings in order
to generate models. The approach presented here requires no
modifications of the application, no architectural abstractions, nor
any structural descriptions of the input data being modeled. The
techniques used are purely mathematical and only need performance data
as input.

Among the statistical models presented in prior works,
\cite{bailey2005performance} specifically mention that it is difficult
for the simplified models to capture variability introduced by
I/O. System variability in general has become a problem of increasing
interest to the HPC and systems communities, however most of the work
has focused on operating system (OS) induced variability
\cite{beckman2008benchmarking,de2007identifying}. The work that has
focused on managing I/O variability does not use any sophisticated
modeling techniques \cite{lofstead2010managing}. Hence, Chapter
\ref{ch:naive} presents a case study applying advanced mathematical
and statistical modeling techniques to the domain of HPC I/O
characteristics. The models are used to predict the mean throughput of
a system and the variance in throughput of a system. The discussion
section that follows outlines how the techniques presented can be
applied to any performance metric and any system.

>>>>

Many previous works attempting to model system performance have used simulated environments \cite{grobelny2007fase,wang2009simulation,wang2013towards}. \citet{grobelny2007fase} refer to statistical models as being oversimplified and not capable of capturing the true complexity of the underlying system. Historically, statistical models have been limited to modelling at most one or two system parameters and have therefore not been capable of modeling the complexity of the underlying system \cite{snavely2002framework,bailey2005performance,barker2009using,ye2010analyzing}. These limited statistical models have provided satisfactory performance in narrow application settings. However these techniques require additional code annotations, hardware abstractions, or additional assumptions about the behavior of the application in order to generate models. In contrast, the approaches that are presented here require no modifications of applications, no architectural abstractions, nor any structural descriptions of the input data being modeled. The techniques used in this work only require performance data as input.

@@

Regression and interpolation have a considerable theoretical base in
one dimension \cite{cheney2009course}. Splines in particular are well
understood as an interpolation technique in one dimension
\cite{de1978practical}, particularly B-splines. Tensor products of
B-splines \cite{unther1996interpolating} or other basis functions have
an unfortunate exponential scaling with increasing
dimension. Exponential scaling prohibits tensor products from being
reasonably applied beyond three-dimensional data. In order to address
this dimensional scaling challenge, C. de Boor and others proposed box
splines \cite{de2013box}, of which one of the approximation techniques
in Chapter \ref{ch:boxes} of this work is composed \cite{lux2018novel}.

@@

Among the statistical models presented in prior works for modeling computer systems,
\cite{bailey2005performance} specifically mention that it is difficult
for simplified models to capture variability introduced by
I/O. System variability in general has become a problem of increasing
interest to the HPC and systems communities, however most of the work
has focused on operating system (OS) induced variability
\cite{beckman2008benchmarking,de2007identifying}. The work that has
focused on managing I/O variability does not use any sophisticated
modeling techniques \cite{lofstead2010managing}. Hence, Chapter
\ref{ch:naive} presents a case study applying advanced mathematical
and statistical modeling techniques to the domain of HPC I/O
characteristics. The models are used to predict the mean throughput of
a system and the variance in throughput of a system. The discussion
section that follows outlines how the techniques presented can be
applied to any performance metric and any system.

### 2020-07-28 3-Disseration/boxes.tex
Updated thesis document, sent draft to Dr. Watson.

>>>>

\begin{figure}
  \centering
  \includegraphics[width=0.6\textwidth,]{Figures/NA/VM_Construction.pdf}
  \caption{Above is a depiction of the Voronoi cell
    boundaries (dashed lines) about a set of interpolation points
    (dots) in two dimensions. In this example, the Voronoi mesh basis
    function about the center most point has nonzero weight in the
    shaded region and transitions from a value of one at the point to
    zero at the boundary of the twice expanded Voronoi cell (solid
    line).
  \vspace{-.1cm}}
  \label{fig:voronoi-basis}
\end{figure}
